Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-02-23 doc/manuscript/ESSDmanuscript.Rmd
minor formatting

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'essd')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "5" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "6" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,5 *"
    email: teixeirak@si.edu
    corresponding: true
affiliation:
  - code: 1
    address: Center for Conservatiton Ecology, Smithsonian Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 3
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 4
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 5
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 6
    address: CIRAD, Montpellier, France
abstract: |
  Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
  Yet, the efficacy of greenhouse gas (GHG) mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
  Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
  Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
  Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector GHG emissions and removals.
  We begin by reconciling terminology and mapping ForC fields into EFDB.
  This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
  At the time of writing, ForC contained ## values that would qualify for inclusion in the EFDB, ## of which have been transferred to date.
  (Some analysis of representation/ gaps)
  <!-- -->
  In the future, forest C estimates in EFDB can be improved through targetted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.

@@

The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.

>>>>

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'essd')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "5" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "6" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,5 *"
    email: teixeirak@si.edu
    corresponding: true
affiliation:
  - code: 1
    address: Center for Conservatiton Ecology, Smithsonian Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 3
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 4
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 5
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 6
    address: CIRAD, Montpellier, France
abstract: |
  Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
  Yet, the efficacy of greenhouse gas (GHG) mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
  Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
  Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
  Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector GHG emissions and removals.
  We begin by reconciling terminology and mapping ForC fields into EFDB.
  This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
  At the time of writing, ForC contained ## values that would qualify for inclusion in the EFDB, ## of which have been transferred to date.
  (Some analysis of representation/ gaps.)
  <!-- -->
  In the future, forest C estimates in EFDB can be improved through targetted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.

@@

<!--The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.-->

### 2022-02-23 doc/manuscript/ESSDmanuscript.tex
minor formatting

\begin{abstract}
Forests are critical for climate change mitigation and consitute a
substantial portion of planned emissions reductions under the 2015 Paris
Agreement. Yet, the efficacy of greenhouse gas (GHG) mitigation planning
and reporting is dependent upon the quality of available emission
factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published,
yet are not readily accesible to the practitioners compiling national
greenhouse gas inventories. Many of these data have, however, been
compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector GHG emissions and
removals. We begin by reconciling terminology and mapping ForC fields
into EFDB. This process required some updates to the ForC database
structure, leading to the release of a new version of ForC (v4.0;
described here). At the time of writing, ForC contained \#\# values that
would qualify for inclusion in the EFDB, \#\# of which have been
transferred to date. (Some analysis of representation/ gaps) In the
future, forest C estimates in EFDB can be improved through targetted
research to fill critical gaps, reporting of information required by
IPCC, and continued submission of data from scientific publications to
the EFDB.
\end{abstract}

@@

The conclusion goes here. You can modify the section name with
\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.

>>>>

\begin{abstract}
Forests are critical for climate change mitigation and consitute a
substantial portion of planned emissions reductions under the 2015 Paris
Agreement. Yet, the efficacy of greenhouse gas (GHG) mitigation planning
and reporting is dependent upon the quality of available emission
factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published,
yet are not readily accesible to the practitioners compiling national
greenhouse gas inventories. Many of these data have, however, been
compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector GHG emissions and
removals. We begin by reconciling terminology and mapping ForC fields
into EFDB. This process required some updates to the ForC database
structure, leading to the release of a new version of ForC (v4.0;
described here). At the time of writing, ForC contained \#\# values that
would qualify for inclusion in the EFDB, \#\# of which have been
transferred to date. (Some analysis of representation/ gaps.) In the
future, forest C estimates in EFDB can be improved through targetted
research to fill critical gaps, reporting of information required by
IPCC, and continued submission of data from scientific publications to
the EFDB.
\end{abstract}

@@

\newpage
